---
title: "Intro to bulletchartr"
author: "Ryo Nakagawara & Amit Kohli"
output: 
  html_document:
    keep_md: true
        
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(bulletchartr)

df <- readxl::read_xlsx("data/Indicators_Targets_full.xlsx", sheet = "Sheet1")

```

# What is a bullet chart?

The __bullet chart__ was invented by [Stephen Few](https://www.perceptualedge.com/articles/misc/Bullet_Graph_Design_Spec.pdf), for the purpose of showing tons of info in a condensed form in KPIs.

This type of graph is a variation on a typical bar graph with a thick line presenting an important point for that indicator (benchmark, performance target, etc.) and other bars in the background that can signify different levels of performance (low-high, bad-good, etc.). The bullet chart makes it very easy to compare between related measures (e.g. present status versus status at similar time in the past).

The output of the `bullet_chart()` function most closely resembles Stephen Few's design: 

```{r, fig.width=6,fig.height=5, echo=TRUE}
bullet_chart(file_name ="data/Indicators_Targets_ext.xlsx")
```

The single black bar represents the current value of the indicator while the different hue columns represent last week's value (darker hue) and last year's value (lighter hue). The bar for each Indicator show the progression along the horizontal-axis presenting the percentage of the yearly target completed. This axis also shows the percent of the year gone by with the vertical line indicating what exact percentage "Today" is, along this percentage. 


As you can see, the bars show the progression along the horizontal-axis presenting the percentage of the yearly target completed. Also, along this axis is the percent of the year gone by with a vertical line indicating what exact percentage __"Today"__ is along this percentage. It is necessary to use percentages as we have multiple indicators of varying units/parameters for each project! 


### GREYSCALE
Each bar is colored "green" if the Indicator completion is close to or past "Today", "orange" if it's close, and "red" if it is far off track.


```{r, echo=FALSE}
df %>% select(indicator_name, actual, actual_lastweek, actual_lastyear, target) %>% glimpse()
```

This package is based on visualizing M&E deliverables or "Indicators", however, it can be handy for anyone that depends on monitoring Key Performance Indicators (KPIs) or needs to track progress against different targets. 

To use the functions included in this package, one needs to provide an Excel (.xlsx) file containing these exact columns (in the order of appearance): `indicator_name`, `actual`, `actual_lastweek`, `actual_lastyear`, and `Target`. The following sections will describe these variables, as well as the extra variables calculated within the function in more detail. In later versions we hope to streamline this process to make it easier to use. 

* `indicator_name`: the name of the indicator or KPI that you are measuring
* `actual`: the value of the indicator at the current time of viewing ("Today")
* `actual_lastweek`: Last week's value of the indicator
* `actual_lastyear`: Last year's value of the indicator
* `Target`: the target value for the indicator (used to calculate the percent variables)

The percentages along the horizontal axis are calculated by: 

* `Perc`: Value of indicator as percent of yearly taget and percent of the year at the current time
* `PercWeek`: Last week's value of the indicator as percent of yearly target and percent of the year
* `PercYear`: Last year's value of the indicator as percent of yearly target and percent of the year

`BehindBy` is calculated by: `Perc - PercentTime` and shows how far behind the current value of the indicator is to the target value for the current time and shows up as the text above each bar: 

* "OK!": Shows that the current value of the indicator meets the target value for the current time
* "Need ___ more": Shows exactly how much more of the indicator is needed to reach the target value for the current time


## bullet_chart() functions!

Along with the `bullet_chart()` function that you saw above we also have `bullet_chart2()`.
This is similar to the standard bullet chart but uses different thicknesses for the bars as the benchmarks for previous time points (last week and last year) to further
accentuate the difference graphically.

```{r, fig.width=6, fig.height=5, echo=TRUE}
bullet_chart_wide(file_name = "data/Indicators_Targets_full.xlsx")
```


`bullet_chart3()` shows a version with different symbols representing the indicator value for 
last week (diamond) and last year (circle).

```{r,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = "data/Indicators_Targets_full.xlsx")
```


`bullet_chart4()` provides a version with a single colored bar representing the current value
for the indicator along with a black vertical line representing the indicator value at this time
last year.

```{r,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_vline(file_name = "data/Indicators_Targets_full.xlsx")

```

## Different calendars

Today shown on a Fiscal Year calendar (considering that today's physical date is `r Sys.Date()`.)

```{r,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = "data/Indicators_Targets_full.xlsx", cal_type = "fis")
```

... or on a calendar year

```{r,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = "data/Indicators_Targets_full.xlsx", cal_type = "cal")
```

... or using a custom date fed directly into the plotting function.

```{r,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = "data/Indicators_Targets_full.xlsx", cal_type = "2018/02/15")
```


## small version

...

## Future Direction

Currently this package is geared more toward non-R using M&E people (therefore, the Excel file input), however as we develop this package further we want to go towards being able to take different file inputs and make the bullet_chart functions more customizable for general use cases. 
