---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

```{r pkgs, include=FALSE}
library(readxl)
library(dplyr)
library(lubridate)
library(ggplot2)
library(bulletchartr)
```

# bulletchartr

<!-- badges: start -->
  [![Codecov test coverage](https://codecov.io/gh/ACDIVOCATech/bulletchartr/branch/master/graph/badge.svg)](https://codecov.io/gh/ACDIVOCATech/bulletchartr?branch=master)
  [![Travis-CI Build Status](https://travis-ci.org/ACDIVOCATech/bulletchartr.svg?branch=master)](https://travis-ci.org/ACDIVOCATech/bulletchartr)
<!-- badges: end -->

This package, `bulletchartr` is based on visualizing M&E deliverables or "Indicators", however, it can be handy for anyone that depends on monitoring Key Performance Indicators (KPIs) or needs to track progress against different targets.

## Installation

```{r install, eval = FALSE}
# Install the package from GitHub:
# install.packages("devtools")

devtools::install_github("ACDIVOCATech/bulletchartr")
```

The __bullet chart__ was invented by [Stephen Few](https://www.perceptualedge.com/articles/misc/Bullet_Graph_Design_Spec.pIndicatorData), for the purpose of showing tons of info in a condensed form in KPIs.

The output of the `bullet_chart()` function most closely resembles Stephen Few's design: 

```{r bulletchart, fig.width=6,fig.height=8, echo=TRUE}
## load example data
load(read_example("df_bc.rda"))

bullet_chart(dataframe = df_bc)
```

The outputs of `bullet_chart_symbols()`, `bullet_chart_wide()`, and `bullet_chart_vline()` have a different x-axis scale to a regular bullet chart. 

The x-axis represents both the percentage of the yearly target AND the percentage of the year that has passed. There is a vertical line showing `TODAY`, which shows at what percentage of the year __and__ what percentage of the target we are at right now. 
The color inside the bar is __green__ if we are near or past the `TODAY` line, __orange__ when we’re close and __red__ when we're very behind schedule/target.

```{r}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_ext.xlsx"))
```

```{r}
bullet_chart_wide(file_name = read_example("Indicators_Targets_ext.xlsx"))
```

```{r}
bullet_chart_vline(file_name = read_example("Indicators_Targets_ext.xlsx"))
```

Please read the package vignette, "Introduction to bullet charts" for a more detailed overview!

## Future direction

Currently this package is geared more toward non-R using M&E people (therefore, the Excel file input alongside a dataframe input), however as we develop this package further we want to go towards being able to make the `bullet_chart` functions more customizable for general use cases. 
