---
title: "bullchartr functions and arguments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bullchartr functions and arguments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bulletchartr)
```


# {bulletchartr} functions!

## `bullet_chart()`

This is the original chart created by Stephen Few.


```{r bulletchart, fig.width=6,fig.height=5, echo=TRUE}
#bullet_chart(file_name = read_example("Indicators_Targets_ext.xlsx"))
```


## Time-Constraint bullet charts

The following versions are a different type of bullet chart that we use at ACDI/VOCA. The main feature is the standardized scale for progress towards a "Yearly Target" and "Year" expressed as percentages.

To read the chart you need to: 

## `bullet_chart_wide()`

`bullet_chart_wide()` is similar to the standard bullet chart but uses different thicknesses for the bars as the benchmarks for previous time points (last week and last year) to further
accentuate the difference graphically.

```{r wide, fig.width=6, fig.height=5, echo=TRUE}
bullet_chart_wide(file_name = read_example("Indicators_Targets_full.xlsx"),
                  legend = TRUE)
```

## `bullet_chart_symbols()`

`bullet_chart_symbols()` shows a version with different symbols representing the indicator value for 
last week (diamond) and last year (circle).

```{r symbols,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_full.xlsx"))
```

## `bullet_chart_vline()`

`bullet_chart_vline()` provides a version with a single colored bar representing the current value
for the indicator along with a black vertical line representing the indicator value at this time
last year.

```{r vline,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_vline(file_name = read_example("Indicators_Targets_ext.xlsx"),
                   cal_type = "cal")
```


# Function arguments

## Different calendars

The vertical line showing `Today` can be customized depending on whether you are measuring by a fiscal year, a calendar year, or a custom date. `Today` shown on a Fiscal Year calendar (considering that today's physical date is `r Sys.Date()`.) is show below:

```{r cal-type,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_full.xlsx"), 
                     cal_type = "fis")
```

or for a calendar year:

```{r cal-type2,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_wide(file_name = read_example("Indicators_Targets_ext.xlsx"), 
                     cal_type = "cal",
                  legend = TRUE)
```

or using a custom date that you can feed directly into the plotting function:

```{r cal-custom,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_ext.xlsx"), 
                     cal_type = "2018/02/15")
```

By doing this the function will automatically calculate your progress and targets according to the calendar type that you specified.

## `small` version

If you want to see a small version of your "Plot" panel just specify `small` = __TRUE__. This will allow you to quickly check the entire plot without having to enlarge it over and over again in the pop-up window. The small version also hides the text so as to not clutter up the limited space.

```{r small,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_wide(file_name = read_example("Indicators_Targets.xlsx"), 
                  small = "TRUE")
```

## Show legends

You have the option to show legends for both the indicator schedule and the symbols by specifying `legend` = **TRUE**. Default is **FALSE**.

```{r legends,fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets.xlsx"), 
                     legend = TRUE)
```

## Show `Last Week` & `Last Year` info text

```{r}

```


## remove indicators with Targets == NA or 0

```{r}

```


