---
title: "bulletchartr function arguments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bulletchartr function arguments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>"
)
```

```{r setup}
library(bulletchartr)
```

# Function arguments

## Different calendars

The vertical line showing `Today` can be customized depending on whether you are measuring by a fiscal year, a calendar year, or a custom date. `Today` shown on a Fiscal Year calendar (considering that today's physical date is `r Sys.Date()`.) is show below:

```{r cal-type,fig.width=6,fig.height=5, warning=FALSE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_ext.xlsx"), 
                     cal_type = "fis")
```

or for a calendar year:

```{r cal-type2,fig.width=6,fig.height=5, warning=FALSE}
bullet_chart_wide(file_name = read_example("Indicators_Targets_ext.xlsx"), 
                     cal_type = "cal",
                  legend = TRUE)
```

or using a custom date that you can feed directly into the plotting function:

```{r cal-custom,fig.width=6,fig.height=5, warning=FALSE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets_ext.xlsx"), 
                     cal_type = "2019/02/15")
```

By doing this the function will automatically calculate your progress and targets according to the calendar type that you specified.

## `small` version

If you want to see a small version of your "Plot" panel just specify `small` = __TRUE__. This will allow you to quickly check the entire plot without having to enlarge it over and over again in the pop-up window. The small version also hides the text so as to not clutter up the limited space.

```{r small}
bullet_chart_wide(file_name = read_example("Indicators_Targets.xlsx"), 
                  small = TRUE)
```

Do note that by default the informational text will not show up by default when `small = TRUE` even when you set the argument `show_text` to TRUE, you will get a short warning when you do.

## Show legends

You have the option to show legends for both the indicator schedule and the symbols by specifying `legend` = **TRUE**. Default is **FALSE**.

```{r legends, warning=FALSE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets.xlsx"), 
                     legend = TRUE)
```

## Show `Last Week` & `Last Year` info text

```{r, warning=FALSE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets.xlsx"), 
                     legend = TRUE,
                     show_text = TRUE)
```

## Chart type: Static vs. Interactive

An interactive version where you can hover your mouse over the bar to see the informational text is available with the `chart_type` argument. This is because the `show_text` option can make the chart look rather busy at times. The interactivity is handled by the {ggiraph} package.

```{r, warning=FALSE}
bullet_chart_symbols(file_name = read_example("Indicators_Targets.xlsx"), 
                     legend = TRUE,
                     chart_type = "interactive")
```

