---
title: "Introduction to bulletcharts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to bulletcharts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What is a bullet chart?

The __bullet chart__ was invented by [Stephen Few](https://www.perceptualedge.com/articles/misc/Bullet_Graph_Design_Spec.pdf), for the purpose of showing tons of info in a condensed form in KPIs.

This type of graph is a variation on a typical bar graph with a thick line presenting an important point for that indicator (benchmark, performance target, etc.) and other bars in the background that can signify different levels of performance (low-high, bad-good, etc.). The bullet chart makes it very easy to compare between related measures (e.g. present status versus status at similar time in the past).

The output of the `bullet_chart()` function most closely resembles Stephen Few's design: 

```{r bulletchart, fig.width=6,fig.height=5, echo=TRUE}
library(bulletchartr)
#bullet_chart(file_name = read_example("Indicators_Targets_ext.xlsx"))
```

The single black bar represents the current value of the indicator while the different hue columns represent last week's value (darker hue) and last year's value (lighter hue). The bar for each Indicator show the progression along the horizontal-axis presenting the percentage of the yearly target completed. This axis also shows the percent of the year gone by with the vertical line indicating what exact percentage "Today" is, along this percentage. 


## Time-constraint bullet chart

```{r bulletchart-time, fig.width=6,fig.height=5, echo=TRUE}
bullet_chart_wide(file_name = read_example("Indicators_Targets_ext.xlsx"))
```

As you can see, the bars show the progression along the horizontal-axis presenting the percentage of the yearly target completed. Also, along this axis is the percent of the year gone by with a vertical line indicating what exact percentage __"Today"__ is along this percentage. It is necessary to use percentages as we have multiple indicators of varying units/parameters for each project! 

The different grey colored bars represent the values of the indicator at "Last Week" and "Last Year". The grey scaled bars can represent any qualitative ranges such as "bad - good - excellent" or "disabled - repairing - fixed", etc. In the near future we will look to expand the capabilities of this package to allow users to specify these qualitative ranges to fit their needs. 

This package, `bulletchartr` is based on visualizing M&E deliverables or "Indicators", however, it can be handy for anyone that depends on monitoring Key Performance Indicators (KPIs) or needs to track progress against different targets. 

# Resources

* [Bullet Graph Design Specification](https://www.perceptualedge.com/articles/misc/Bullet_Graph_Design_Spec.pdf)
